Ionospheric Disturbances in Mexican Territory Produced by Objects Entering the Athmosphere from Space \tnoteref{tnote1}

### 2021-10-04 paper/methodology.tex
A huge commit:
- Add program to plot Rate of TEC (sTEC or vTEC)
- Improved Program of vTEc maps to plot sTEC maps
- Program to plot Dst index
- Write something in the introduction section
- Improved stations table in methodology section

\subsection{Meteors Database}

We selected a sample of meteors which were observed in mexican territory from the Geostationary Lightning Mapper \citep{GOODMAN:2013}. Orignally this project was designed to detect ligthning activity in earth's athmosphere, but has been proven that also can detect bolides entering the athmosphere. The detection comes from two satellites called GOES-16 and GOES-17 orbiting the earth in geostationary orbits. %Figure  shows the area where GOES-16 and GOES-17 detect activity on a lightning flash rate map; in mexican territory clearly both satellites can make detections. %Create a figure where the area coverage of GOES-16 and GOES-17 is shown.
We used the interactive database available at \url{https://neo-bolide.ndc.nasa.gov/#/}. These data are publicly available and easily downloaded from the same website. For each event we can obtain the recorded trajectory of meteors and the corresponding light curve. THe GLM satellites have an umbral magnitude for detection of -14. At this magnitude, a meteor is considered a bolide, and is expected to be at least decimeter-sized (in diameter) to reach such brightness. In the other hand, too bright meteors will saturate the detectors, and thus, lowering the quality of data. The result of this factors implies that the range in size of the objects in our sample varies in diameter between decimeter to meter size. %The sample was chosen following the next criteria:%selected in such way we chose the most probable objects to be detected by GPS sources in mexican territory and its surroundings: 
Each event also has assigned a confidence ratio, from low confidence to high, depending in how bright is the event itself and if the trajectory recorded by GLM ressembles (or not) a straight line. We chose only events whose confidence ratio is high, in oreder to be sure we chose the brightest objects, and thus, in the diameter size of bolides, we favored the meter-sized ones. In table \ref{tab:table-meteors} we list the object we chose to do this work, order in chronological order. The columns of the table, from left to right are and ID to enumerate the meteors in the sample, the date and time each meteor was detected, the duration of the detection, their respective coordinates and the estimated height of the meteor over the ground at the time of the detection. GOES-16 and GOES-17 systematically detect the meteors at slightly different positions and at slightly different times, so we calculated the mean of the duration, latitude and longitude reported by both satellites for each event, and used the standard deviation as the uncertainties.

@@

\begin{table*}
  \centering
  \caption{List of meteors passing through Mexico. The events are listed in chronological order. The listed duration, latitude and longitude correspond to the mean of the measurements of both GOES satellites. The uncertainties correspond to the respecting mean deviation.}
\label{tab:table-meteors}
\begin{tabular}{rrrrrrr}
\hline
ID & Date of event & Start Time (UT)  & Duration (seconds) & Latitude (deg) & Longitude (deg) & Height (km)\\
\hline
01 & 2019-05-23 & 16:36:18 & $0.197\pm 0.0000$ & $24.30 \pm 0.000$ & $-101.60 \pm 0.849$  & 28\\
02 & 2019-07-18 & 14:30:30 & $0.058\pm 0.0000$ & $27.20 \pm 0.000$ & $-103.15 \pm 0.778$  & 72\\
03 & 2019-08-10 & 11:18:48 & $0.199\pm 0.0757$ & $21.50 \pm 0.000$ & $-102.50  \pm 0.849$ & 92\\
04 & 2019-10-03 & 07:55:33 & $0.106\pm 0.0297$ & $25.65 \pm 0.071$ & $-96.25 \pm   0.778$ & 74\\
05 & 2019-10-09 & 06:08:11 & $0.103\pm 0.0078$ & $23.60 \pm 0.000$ & $-111.95 \pm  0.212$ & 32\\
06 & 2019-11-16 & 09:36:04 & $0.396\pm 0.0134$ & $20.30 \pm 0.000$ & $-100.55 \pm  0.919$ & 82\\
07 & 2019-11-17 & 15:36:01 & $0.116\pm 0.0035$ & $31.70 \pm 0.000$ & $-117.70 \pm  1.131$ & 88\\
08 & 2019-11-19 & 07:57:40 & $0.097\pm 0.1138$ & $20.00 \pm 0.000$ & $-88.40 \pm  1.131$  & 99\\
09 & 2019-11-26 & 13:23:20 & $0.078\pm 0.0290$ & $23.90 \pm 0.000$ & $-108.70 \pm  0.849$ & 81\\
10 & 2019-12-04 & 09:42:54 & $0.173\pm 0.0028$ & $31.50 \pm 0.000$ & $-113.65 \pm  0.919$ & 77\\
11 & 2019-12-15 & 14:50:49 & $0.127\pm 0.0134$ & $27.70 \pm 0.000$ & $-114.10 \pm  0.849$ & 78\\
12 & 2019-12-29 & 16:16:35 & $0.062\pm 0.0134$ & $29.60 \pm 0.000$ & $-116.35 \pm  0.919$ & 79\\
13 & 2020-01-03 & 14:10:17 & $0.113\pm 0.0085$ & $30.20 \pm 0.000$ & $-117.65 \pm  0.919$ & 74\\
14 & 2020-01-06 & 16:39:27 & $0.118\pm 0.0042$ & $31.40 \pm 0.000$ & $-108.20 \pm  0.990$ & 81\\
15 & 2020-01-15 & 15:00:33 & $0.213\pm 0.1351$ & $19.45 \pm 0.071$ & $-95.55 \pm   0.919$ & 93\\
16 & 2020-02-12 & 09:25:40 & $0.210\pm 0.0226$ & $18.90 \pm 0.000$ & $-93.50 \pm   0.849$ & 90\\
17 & 2020-03-03 & 12:33:27 & $0.062\pm 0.0007$ & $18.25 \pm 0.071$ & $-106.35 \pm 0.636$  & 77\\
18 & 2020-03-31 & 19:31:52 & $0.105\pm 0.0573$ & $28.45 \pm 0.071$ & $-112.05 \pm  0.636$ & 61\\
19 & 2020-04-08 & 16:25:28 & $0.120\pm 0.0926$ & $26.10 \pm 0.000$ & $-93.90 \pm   0.849$ & 78\\
20 & 2020-04-18 & 17:43:25 & $0.139\pm 0.0106$ & $29.00 \pm 0.000$ & $-106.55 \pm  0.919$ & 82\\
21 & 2020-04-20 & 16:05:22 & $0.318\pm 0.1655$ & $28.15 \pm 0.071$ & $-97.85 \pm   1.061$ & 88\\
22 & 2020-04-25 & 11:03:09 & $0.323\pm 0.0813$ & $32.15 \pm 0.071$ & $-111.60 \pm  1.131$ & 84\\
23 & 2020-04-28 & 19:31:52 & $0.105\pm 0.0573$ & $28.45 \pm 0.071$ & $-112.05 \pm  0.636$ & 29\\
24 & 2020-05-08 & 10:06:16 & $0.490\pm 0.0750$ & $21.60 \pm 0.000$ & $-92.40 \pm   0.849$ & 81\\
25 & 2020-07-15 & 19:58:28 & $0.693\pm 0.0495$ & $24.00 \pm 0.000$ & $-108.35 \pm  0.495$ & 53\\
26 & 2020-08-07 & 13:29:57 & $0.163\pm 0.0057$ & $28.80 \pm 0.000$ & $-106.05 \pm  0.919$ & 89\\
27 & 2020-09-13 & 16:41:59 & $0.184\pm 0.0078$ & $28.45 \pm 0.071$ & $-113.75 \pm  0.919$ & 85\\
28 & 2020-09-30 & 12:28:11 & $0.100\pm 0.0078$ & $24.90 \pm 0.000$ & $-110.90 \pm  0.849$ & 83\\
29 & 2020-11-16 & 12:28:11 & $0.100\pm 0.0078$ & $24.90 \pm 0.000$ & $-110.90 \pm  0.849$ & 06\\
30 & 2020-11-17 & 12:53:41 & $0.404\pm 0.0262$ & $23.00 \pm 0.000$ & $-102.45 \pm  0.919$ & 93\\
31 & 2020-12-19 & 10:18:14 & $0.407\pm 0.0110$ & $21.95 \pm 0.071$ & $-101.60 \pm  0.990$ & 98\\
32 & 2020-12-23 & 09:43:01 & $0.148\pm 0.0014$ & $25.75 \pm 0.071$ & $-111.25 \pm  0.778$ & 81\\
33 & 2020-12-29 & 15:20:54 & $0.118\pm 0.0014$ & $16.80 \pm 0.000$ & $-102.20 \pm  0.707$ & 81\\
34 & 2021-03-31 & 09:01:17 & $0.753\pm 0.3083$ & $20.15 \pm 0.071$ & $-92.95 \pm  0.212$  & 24\\\hline
\end{tabular}
\end{table*}

@@

\subsection{GPS data}

This material is based on services provided by the GAGE Facility, operated by UNAVCO, Inc., with support from the National Science Foundation and the National Aeronautics and Space Administration under NSF Cooperative Agreement EAR-1724794.

We got RINEX data from 3 to 7 stations depending of the event location and data availability that surround the event place in all directions as possible. A list of the stations where we got RINEX data is available in table  . Most of the stations lie in mexican territory, but in some cases we required data from other stations to cover events near the mexican frontier at north or south.

@@

  \begin{longtable}{lllp{4cm}p{10cm}}

@@

    Station name & Latitude & Longitude & Date of events  & Citation \\\hline
    BAR1\hyperlink{Hudnut}{${}^1$}\hyperlink{Hudnut2}{${}^5$}& 33.48 & -119.03 & 2019-12-29 2020-01-03 & UNAVCO Community, Hudnut, Kenneth, King, Nancy, Aspiotes, Aris G., Borsa, Adrian A., Determan, Daniel N., Galetzka, John E., Stark, Keith F., 2005, SCIGN-PBO Nucleus GPS Network - BAR1-Santa Barbara Island One P.S., The GAGE Facility operated by UNAVCO, Inc., GPS/GNSS Observations Dataset, \url{https://doi.org/10.7283/T5668BHN}.\\
    BLYT\hyperlink{Hudnut}{${}^1$} & 33.61 & -114.71 & 2019-12-29 2020-01-03 & Hudnut, Kenneth, King, Nancy, Aspiotes, Aris G., Borsa, Adrian A., Determan, Daniel N., Galetzka, John E., Stark, Keith F., 2006, SCIGN USGS GPS Network - BLYT-Blythe P.S., The GAGE Facility operated by UNAVCO, Inc., GPS/GNSS Observations Dataset, \url{https://doi.org/10.7283/T5HT2MKK}.\\
    CN23 & 17.26 & -88.78 & 2019-11-19 2020-01-15 2020-02-12& UNAVCO Community, 2012, COCONet GPS Network - CN23-BelmopanBZCR2012 P.S., The GAGE Facility operated by UNAVCO, Inc., GPS/GNSS Observations Dataset, \url{https://doi.org/10.7283/T5Q23XJH}.\\
    CN25 & 16.23 & -92.13 & 2020-01-15 & UNAVCO Community, 2014, COCONet GPS Network - CN25-ComitandDMEX2012 P.S., The GAGE Facility operated by UNAVCO, Inc., GPS/GNSS Observations Dataset, \url{https://doi.org/10.7283/T57W69G7}.\\
    GCFS & 19.31 & -81.18 & 2019-11-19& Watts, Anthony, 2016, COCONet GPS Network - GCFS-G\_CAYMAN\_CYM2014 P.S., The GAGE Facility operated by UNAVCO, Inc., GPS/GNSS Observations Dataset, \url{https://doi.org/10.7283/7ETV-X536}.\\
    GMPK\hyperlink{Hudnut}{${}^1$} & 33.05 & -114.83 & 2019-12-04 & UNAVCO Community, Hudnut, Kenneth, King, Nancy, Aspiotes, Aris G., Borsa, Adrian A., Determan, Daniel N., Galetzka, John E., Stark, Keith F., 2005, SCIGN-PBO Nucleus GPS Network - GMPK-Glamis Peak P.S., The GAGE Facility operated by UNAVCO, Inc., GPS/GNSS Observations Dataset, \url{https://doi.org/10.7283/WCHN-H687}.\\
    GUAT\hyperlink{Garnier}{${}^2$} & 14.59 & -90.52 & 2020-02-12 & DeMets, Charles, Cosenza-Muralles, Beatriz, 2021, Central America 2018 - Guatemala, The GAGE Facility operated by UNAVCO, Inc., GPS/GNSS Observations Dataset, \url{https://doi.org/10.7283/KH2R-K704}.\\
    GUAX\hyperlink{Hudnut}{${}^1$} & 28.88 & -118.29 & 2019-10-09 2019-12-15 2019-12-29 2020-01-03 2020-03-31 2020-07-15 2020-09-13 2020-09-30 2020-12-23 & Hudnut, Kenneth, King, Nancy, Aspiotes, Aris G., Borsa, Adrian A., Determan, Daniel N., Galetzka, John E., Stark, Keith F., 2001, SCIGN USGS GPS Network - GUAX-Isla Guadalupe P.S., The GAGE Facility operated by UNAVCO, Inc., GPS/GNSS Observations Dataset, \url{https://doi.org/10.7283/T5GX48T2}.\\
    IAGX & 29.03 & -113.17 & 2019-12-04 & Gonzalez-Ortega, Alejandro, Galetzka, John E., Gonzalez, Javier, 2018, CICESE REGNOM GPS Network - IAGX-iagxREGNOMmx2018 P.S., The GAGE Facility operated by UNAVCO, Inc., GPS/GNSS Observations Dataset, \url{https://doi.org/10.7283/DGWN-A627}.\\
    INEG & 21.85 & -102.28 & 2020-07-15 2020-08-07 2020-09-30 2020-11-16 2020-11-17 2020-12-19 & No citations were found\\
    KVTX & 27.55 & -97.89 & 2019-05-23 2019-07-18 2019-08-10 2019-10-03 2019-11-17 2020-04-08 2020-04-18 2020-04-20 2020-05-08 2020-08-07 & UNAVCO Community, 2007, PBO GPS Network - KVTX-KingsvilleTX2006 P.S., The GAGE Facility operated by UNAVCO, Inc., GPS/GNSS Observations Dataset, \url{https://doi.org/10.7283/T5J38QH8}.\\
    MDO1 & 30.68 & -104.02 & 2019-07-18& No citations were found\\
    MGO5 & 30.68 & -104.02 & 2020-04-20 2020-08-07 & No citations were found\\
    MGW3 & 29.62 & -89.95 & 2020-04-08 2020-04-20 2020-05-08 & No citations were found\\
    OXTH & 16.29 & -95.24 & 2020-01-15 2020-02-12 & DeMets, Charles, Cabral-Cano, Enrique, 2008, Oaxaca GPS Network - OXTH-Tehuantepec P.S., The GAGE Facility operated by UNAVCO, Inc., GPS/GNSS Observations Dataset, \url{https://doi.org/10.7283/T5Q81B5V}.\\
    OXUM\hyperlink{Graham}{${}^3$} & 15.66 & -96.50 & 2021-03-31 & Cabral-Cano, Enrique, Salazar-Tlaczani, Luis, 2015, TLALOCNet - OXUM-oxum\_tnet\_mx2001 P.S., The GAGE Facility operated by UNAVCO, Inc., GPS/GNSS Observations Dataset, \url{https://doi.org/10.7283/T5J964RP}.\\
    P001 & 31.95 & -112.80 & 2020-04 25 & UNAVCO Community, 2008, PBO GPS Network - P001-Organ\_PipeAZ2007 P.S., The GAGE Facility operated by UNAVCO, Inc., GPS/GNSS Observations Dataset, \url{https://doi.org/10.7283/T5DR2SGP}.\\
    P014 & 31.97 & -11.09 & 2019-12-04 2019-12-29 2020-01-03 2020-01-06 2020-04-25 & UNAVCO Community, 2008, PBO GPS Network - P014-Sahuarita\_AZ2007 P.S., The GAGE Facility operated by UNAVCO, Inc., GPS/GNSS Observations Dataset, \url{https://doi.org/10.7283/T5DJ5CMK}.\\
    P807 & 30.49 & -98.82 & 2019-11-17 2020-01-06 2020-04-20 2020-11-17 & UNAVCO Community, 2012, PBO GPS Network - P807-EcRockStPkTX2012 P.S., The GAGE Facility operated by UNAVCO, Inc., GPS/GNSS Observations Dataset, \url{https://doi.org/10.7283/T5TQ5ZKM}.\\
    PLPX & 31.59 & -115.15 & 2019-12-04 & UNAVCO Community, 2011, PBO GPS Network - PLPX-Las\_PintasMX2010 P.S., The GAGE Facility operated by UNAVCO, Inc., GPS/GNSS Observations Dataset, \url{https://doi.org/10.7283/T5K64G3T}.\\
    PTEX & 32.29 & -116.52 & 2019-12-29 2020-01-03 2020-09-13 2020-12-23 & UNAVCO Community, 2011, PBO GPS Network - PTEX-Testerazo\_MX2011 P.S., The GAGE Facility operated by UNAVCO, Inc., GPS/GNSS Observations Dataset, \url{https://doi.org/10.7283/T5610XBP}.\\
    RG06 & 32.63 & -107.86 & 2020-04-25 & Sheehan, Anne, 2007, Rio Grande Rift GPS Network - RG06-RG06FaywodNM2006 P.S., The GAGE Facility operated by UNAVCO, Inc., GPS/GNSS Observations Dataset, \url{https://doi.org/10.7283/T5668BFR}.\\
    RG07 & 32.50 & -106.84 & 2020-01-06 & Sheehan, Anne, 2007, Rio Grande Rift GPS Network - RG07-RG07CrucesNM2006 P.S., The GAGE Facility operated by UNAVCO, Inc., GPS/GNSS Observations Dataset, \url{https://doi.org/10.7283/T5KD1W45}. \\

>>>>

\subsection{Meteors Databases}
\label{ssec:databases}
We selected a sample of meteors which were observed in mexican territory from the Geostationary Lightning Mapper \citep{GOODMAN:2013}. Orignally this project was designed to detect ligthning activity in earth's athmosphere, but has been proven that also can detect bolides entering the athmosphere. The detection comes from two satellites called GOES-16 and GOES-17 orbiting the earth in geostationary orbits. %Figure  shows the area where GOES-16 and GOES-17 detect activity on a lightning flash rate map; in mexican territory clearly both satellites can make detections. %Create a figure where the area coverage of GOES-16 and GOES-17 is shown.
We used the interactive database available at \url{https://neo-bolide.ndc.nasa.gov/#/}. These data are publicly available and easily downloaded from the same website. For each event we can obtain the recorded trajectory of meteors and the corresponding light curve. THe GLM satellites have an umbral magnitude for detection of -14. At this magnitude, a meteor is considered a bolide, and is expected to be at least decimeter-sized (in diameter) to reach such brightness. In the other hand, too bright meteors will saturate the detectors, and thus, lowering the quality of data. The result of this factors implies that the range in size of the objects in our sample varies in diameter between decimeter to meter size. %The sample was chosen following the next criteria:%selected in such way we chose the most probable objects to be detected by GPS sources in mexican territory and its surroundings: 
Each event also has assigned a confidence ratio, from low confidence to high, depending in how bright is the event itself and if the trajectory recorded by GLM ressembles (or not) a straight line. We chose only events whose confidence ratio is high, in oreder to be sure we chose the brightest objects, and thus, in the diameter size of bolides, we favored the meter-sized ones. In table \ref{tab:table-meteors} we list the object we chose to do this work, order in chronological order. The columns of the table, from left to right are and ID to enumerate the meteors in the sample, the date and time each meteor was detected, the duration of the detection, their respective coordinates and the estimated height of the meteor over the ground at the time of the detection. GOES-16 and GOES-17 systematically detect the meteors at slightly different positions and at slightly different times, so we calculated the mean of the duration, latitude and longitude reported by both satellites for each event, and used the standard deviation as the uncertainties.

@@

\begin{table*}
  \centering
  \caption{List of meteors passing through Mexico. The events are listed in chronological order. The listed duration, latitude and longitude correspond to the mean of the measurements of both GOES satellites. The uncertainties correspond to the respecting mean deviation.}
\label{tab:table-meteors}
\begin{tabular}{rrrrrrrr}
ID & Date of event & Start Time (UT)  & Duration (seconds) & Latitude (deg) & Longitude (deg) & Height (km) & Maximum Kp index \\
GLM-01 & 2019-05-23 & 16:36:18 & $0.197\pm 0.0000$ & $24.30 \pm 0.000$ & $-101.60 \pm 0.849$  & 28 & 2\\
GLM-02 & 2019-07-18 & 14:30:30 & $0.058\pm 0.0000$ & $27.20 \pm 0.000$ & $-103.15 \pm 0.778$  & 72 & 1\\
GLM-03 & 2019-08-10 & 11:18:48 & $0.199\pm 0.0757$ & $21.50 \pm 0.000$ & $-102.50  \pm 0.849$ & 92 & 3\\
GLM-04 & 2019-10-03 & 07:55:33 & $0.106\pm 0.0297$ & $25.65 \pm 0.071$ & $-96.25 \pm   0.778$ & 74 & 2\\
GLM-05 & 2019-10-09 & 06:08:11 & $0.103\pm 0.0078$ & $23.60 \pm 0.000$ & $-111.95 \pm  0.212$ & 32 & 4\\
GLM-06 & 2019-11-16 & 09:36:04 & $0.396\pm 0.0134$ & $20.30 \pm 0.000$ & $-100.55 \pm  0.919$ & 82 & 2\\
GLM-07 & 2019-11-17 & 15:36:01 & $0.116\pm 0.0035$ & $31.70 \pm 0.000$ & $-117.70 \pm  1.131$ & 88 & 3\\
GLM-08 & 2019-11-19 & 07:57:40 & $0.097\pm 0.1138$ & $20.00 \pm 0.000$ & $-88.40 \pm  1.131$  & 99 & 1\\
GLM-09 & 2019-11-26 & 13:23:20 & $0.078\pm 0.0290$ & $23.90 \pm 0.000$ & $-108.70 \pm  0.849$ & 81 & 2\\
GLM-10 & 2019-12-04 & 09:42:54 & $0.173\pm 0.0028$ & $31.50 \pm 0.000$ & $-113.65 \pm  0.919$ & 77 & 2\\
GLM-11 & 2019-12-15 & 14:50:49 & $0.127\pm 0.0134$ & $27.70 \pm 0.000$ & $-114.10 \pm  0.849$ & 78 & 2\\
GLM-12 & 2019-12-29 & 16:16:35 & $0.062\pm 0.0134$ & $29.60 \pm 0.000$ & $-116.35 \pm  0.919$ & 79 & 1\\
GLM-13 & 2020-01-03 & 14:10:17 & $0.113\pm 0.0085$ & $30.20 \pm 0.000$ & $-117.65 \pm  0.919$ & 74 & 3\\
GLM-14 & 2020-01-06 & 16:39:27 & $0.118\pm 0.0042$ & $31.40 \pm 0.000$ & $-108.20 \pm  0.990$ & 81 & 4\\
GLM-15 & 2020-01-15 & 15:00:33 & $0.213\pm 0.1351$ & $19.45 \pm 0.071$ & $-95.55 \pm   0.919$ & 93 & 2\\
GLM-16 & 2020-02-12 & 09:25:40 & $0.210\pm 0.0226$ & $18.90 \pm 0.000$ & $-93.50 \pm   0.849$ & 90 & 2\\
GLM-17 & 2020-03-03 & 12:33:27 & $0.062\pm 0.0007$ & $18.25 \pm 0.071$ & $-106.35 \pm 0.636$  & 77 & 2\hyperlink{note2}{${}^2$}\\
GLM-18 & 2020-03-31 & 19:31:52 & $0.105\pm 0.0573$ & $28.45 \pm 0.071$ & $-112.05 \pm  0.636$ & 61 & 4\\
GLM-19 & 2020-04-08 & 16:25:28 & $0.120\pm 0.0926$ & $26.10 \pm 0.000$ & $-93.90 \pm   0.849$ & 78 & 3\\
GLM-20 & 2020-04-18 & 17:43:25 & $0.139\pm 0.0106$ & $29.00 \pm 0.000$ & $-106.55 \pm  0.919$ & 82 & 2\\
GLM-21 & 2020-04-20 & 16:05:22 & $0.318\pm 0.1655$ & $28.15 \pm 0.071$ & $-97.85 \pm   1.061$ & 88 & 5\\
GLM-22 & 2020-04-25 & 11:03:09 & $0.323\pm 0.0813$ & $32.15 \pm 0.071$ & $-111.60 \pm  1.131$ & 84 & 3\\
GLM-23 & 2020-04-28 & 19:31:52 & $0.105\pm 0.0573$ & $28.45 \pm 0.071$ & $-112.05 \pm  0.636$ & 29 & 3\\
GLM-24 & 2020-05-08 & 10:06:16 & $0.490\pm 0.0750$ & $21.60 \pm 0.000$ & $-92.40 \pm   0.849$ & 81 & 1\\
GLM-25 & 2020-07-15 & 19:58:28 & $0.693\pm 0.0495$ & $24.00 \pm 0.000$ & $-108.35 \pm  0.495$ & 53 & 2\hyperlink{note}{${}^1$}\\
GLM-26 & 2020-08-07 & 13:29:57 & $0.163\pm 0.0057$ & $28.80 \pm 0.000$ & $-106.05 \pm  0.919$ & 89 & 2\\
GLM-27 & 2020-09-13 & 16:41:59 & $0.184\pm 0.0078$ & $28.45 \pm 0.071$ & $-113.75 \pm  0.919$ & 85 & 3\\
GLM-28 & 2020-09-30 & 12:28:11 & $0.100\pm 0.0078$ & $24.90 \pm 0.000$ & $-110.90 \pm  0.849$ & 83 & 4\\
GLM-29 & 2020-11-16 & 12:28:11 & $0.100\pm 0.0078$ & $24.90 \pm 0.000$ & $-110.90 \pm  0.849$ & 06 & 1\\
GLM-30 & 2020-11-17 & 12:53:41 & $0.404\pm 0.0262$ & $23.00 \pm 0.000$ & $-102.45 \pm  0.919$ & 93 & 1\\
GLM-31 & 2020-12-19 & 10:18:14 & $0.407\pm 0.0110$ & $21.95 \pm 0.071$ & $-101.60 \pm  0.990$ & 98 & 2\\
GLM-32 & 2020-12-23 & 09:43:01 & $0.148\pm 0.0014$ & $25.75 \pm 0.071$ & $-111.25 \pm  0.778$ & 81 & 4\\
GLM-33 & 2020-12-29 & 15:20:54 & $0.118\pm 0.0014$ & $16.80 \pm 0.000$ & $-102.20 \pm  0.707$ & 81 & 3\\
GLM-34 & 2021-03-31 & 09:01:17 & $0.753\pm 0.3083$ & $20.15 \pm 0.071$ & $-92.95 \pm  0.212$  & 24 & 3\\\hline
\end{tabular}
      \hypertarget{note}{${}^1$} Kp index is 4 the previous day of impact

      \hypertarget{note2}{${}^2$}Kp index is 4 two days before the impact 
    \end{minipage}

\end{table*}

@@

\subsection{GPS data}
\label{ssec:GPS}
This material is based on services provided by the GAGE Facility, operated by UNAVCO, Inc., with support from the National Science Foundation and the National Aeronautics and Space Administration under NSF Cooperative Agreement EAR-1724794.

We got RINEX data from 3 to 7 stations depending of the event location and data availability that surround the event place in all directions as possible. A list of the stations where we got RINEX data is available in table \ref{tab:table-stations}. Most of the stations lie in mexican territory, but in some cases we required data from other stations to cover events near the mexican frontier at north or south.

@@

  \begin{longtable}{llllllp{12cm}}

@@

    Station name & Latitude & Longitude & \multicolumn{3}{c}{Events ID}  & Citation \\\hline
    \multirow{4}{*}{BAR1\hyperlink{Hudnut}{${}^1$}\hyperlink{Hudnut2}{${}^5$}} & \multirow{4}{*}{33.48} & \multirow{4}{*}{-119.03} & \multirow{4}{*}{GLM-07} & \multirow{4}{*}{GLM-12} & \multirow{4}{*}{GLM-13} & UNAVCO Community, Hudnut, Kenneth, King, Nancy, Aspiotes, Aris G., Borsa, Adrian A., Determan, \\
    &&&&&& Daniel N., Galetzka, John E., Stark, Keith F., 2005, SCIGN-PBO Nucleus GPS Network - BAR1-Santa \\
    &&&&&& Barbara Island One P.S., The GAGE Facility operated by UNAVCO, Inc., GPS/GNSS Observations Dataset, \\
    &&&&&& \url{https://doi.org/10.7283/T5668BHN}.\\\hline
    \multirow{3}{*}{BLYT\hyperlink{Hudnut}{${}^1$}} & \multirow{3}{*}{33.61} & \multirow{3}{*}{-114.71} &  \multirow{3}{*}{GLM-12} & \multirow{3}{*}{GLM-13}  & & Hudnut, Kenneth, King, Nancy, Aspiotes, Aris G., Borsa, Adrian A., Determan, Daniel N., Galetzka, John \\
    &&&&&& E., Stark, Keith F., 2006, SCIGN USGS GPS Network - BLYT-Blythe P.S., The GAGE Facility operated \\
    &&&&&&  by UNAVCO, Inc., GPS/GNSS Observations Dataset, \url{https://doi.org/10.7283/T5HT2MKK}.\\\hline
    \multirow{3}{*}{CN23} & \multirow{3}{*}{17.26} & \multirow{3}{*}{-88.78} &\multirow{3}{*}{GLM-08} & \multirow{3}{*}{GLM-15} & \multirow{3}{*}{GLM-16} & UNAVCO Community, 2012, COCONet GPS Network - CN23-BelmopanBZCR2012 P.S., The GAGE \\
    &&&&&& Facility operated by UNAVCO, Inc., GPS/GNSS Observations Dataset, \\
    &&&&&& \url{https://doi.org/10.7283/T5Q23XJH}.\\\hline
    \multirow{3}{*}{CN25} & \multirow{3}{*}{16.23} & \multirow{3}{*}{-92.13} & \multirow{3}{*}{GLM-15} & & & UNAVCO Community, 2014, COCONet GPS Network - CN25-ComitandDMEX2012 P.S., The GAGE Facility operated by UNAVCO, Inc., GPS/GNSS Observations Dataset, \url{https://doi.org/10.7283/T57W69G7}.\\\hline
    \multirow{3}{*}{GCFS} & \multirow{3}{*}{19.31} & \multirow{3}{*}{-81.18} & \multirow{3}{*}{GLM-08} & & & Watts, Anthony, 2016, COCONet GPS Network - GCFS-G\_CAYMAN\_CYM2014 P.S., The GAGE Facility operated by UNAVCO, Inc., GPS/GNSS Observations Dataset, \url{https://doi.org/10.7283/7ETV-X536}.\\\hline
    \multirow{4}{*}{GMPK\hyperlink{Hudnut}{${}^1$}} & \multirow{4}{*}{33.05} & \multirow{4}{*}{-114.83} & \multirow{4}{*}{GLM-10} & & & UNAVCO Community, Hudnut, Kenneth, King, Nancy, Aspiotes, Aris G., Borsa, Adrian A., Determan, \\
    &&&&&& Daniel N., Galetzka, John E., Stark, Keith F., 2005, SCIGN-PBO Nucleus GPS Network - GMPK-Glamis \\
    &&&&&& Peak P.S., The GAGE Facility operated by UNAVCO, Inc., GPS/GNSS Observations Dataset, \\
    &&&&&& \url{https://doi.org/10.7283/WCHN-H687}.\\\hline
    \multirow{3}{*}{GUAT\hyperlink{Garnier}{${}^2$}} & \multirow{3}{*}{14.59} & \multirow{3}{*}{-90.52} & \multirow{3}{*}{GLM-16} & & & DeMets, Charles, Cosenza-Muralles, Beatriz, 2021, Central America 2018 - Guatemala, The GAGE \\
    &&&&&& Facility operated by UNAVCO, Inc., GPS/GNSS Observations Dataset, \\
    &&&&&& \url{https://doi.org/10.7283/KH2R-K704}.\\\hline
\multirow{4}{*}{GUAX\hyperlink{Hudnut}{${}^1$}} & \multirow{4}{*}{28.88} & \multirow{4}{*}{-118.29} & GLM-05 & GLM-07 & GLM-11 & Hudnut, Kenneth, King, Nancy, Aspiotes, Aris G., Borsa, Adrian A., Determan, Daniel N., Galetzka, John \\
    &&&GLM-12& GLM-13& GLM-18 & E., Stark, Keith F., 2001, SCIGN USGS GPS Network - GUAX-Isla Guadalupe P.S., The GAGE Facility \\
    &&& GLM-25 &GLM-27 &GLM-28& operated by UNAVCO, Inc., GPS/GNSS Observations Dataset, \url{https://doi.org/10.7283/T5GX48T2}.\\
    &&& GLM-32 &       &     &  \\\hline
    \multirow{3}{*}{IAGX} & \multirow{3}{*}{29.03} & \multirow{3}{*}{-113.17} & \multirow{3}{*}{GLM-10} & & & Gonzalez-Ortega, Alejandro, Galetzka, John E., Gonzalez, Javier, 2018, CICESE REGNOM GPS Network \\
    &&&&&& - IAGX-iagxREGNOMmx2018 P.S., The GAGE Facility operated by UNAVCO, Inc., GPS/GNSS  \\
    &&&&&& Observations Dataset, \url{https://doi.org/10.7283/DGWN-A627}. \\\hline
    \multirow{2}{*}{INEG} & \multirow{2}{*}{21.85} & \multirow{2}{*}{-102.28} & GLM-25 & GLM-26 & GLM-28  & No citations were found \\
    &&& GLM-29 & GLM-30 & GLM-31 & \\\hline 
    \multirow{4}{*}{KVTX} & \multirow{4}{*}{27.55} & \multirow{4}{*}{-97.89} & GLM-01 & GLM-02 & GLM-03  & UNAVCO Community, 2007, PBO GPS Network - KVTX-KingsvilleTX2006 P.S., The GAGE Facility \\
    &&& GLM-04 & GLM-06 & GLM-19 & operated by UNAVCO, Inc., GPS/GNSS Observations Dataset, \url{https://doi.org/10.7283/T5J38QH8}.\\
    &&& GLM-20 & GLM-21 & GLM-24 & \\
    &&& GLM-26 &&& \\\hline
    MDO1 & 30.68 & -104.02 & GLM-02 & & & No citations were found\\\hline
    MGO5 & 30.68 & -104.02 & GLM-21 &GLM-26 &  & No citations were found\\\hline
    MGW3 & 29.62 & -89.95 & GLM-19 & GLM-21 & GLM-24 & No citations were found\\\hline
    \multirow{3}{*}{OXTH} & \multirow{3}{*}{16.29} & \multirow{3}{*}{-95.24} & \multirow{3}{*}{GLM-15} & \multirow{3}{*}{GLM-16} & & DeMets, Charles, Cabral-Cano, Enrique, 2008, Oaxaca GPS Network - OXTH-Tehuantepec P.S., The \\
    &&&&&& GAGE Facility operated by UNAVCO, Inc., GPS/GNSS Observations Dataset, \\
    &&&&&& \url{https://doi.org/10.7283/T5Q81B5V}.\\\hline
    \multirow{3}{*}{OXUM\hyperlink{Graham}{${}^3$}} & \multirow{3}{*}{15.66} & \multirow{3}{*}{-96.50} & \multirow{3}{*}{GLM-34} & & & Cabral-Cano, Enrique, Salazar-Tlaczani, Luis, 2015, TLALOCNet - OXUM-oxum\_tnet\_mx2001 P.S., The \\
    &&&&&& GAGE Facility operated by UNAVCO, Inc., GPS/GNSS Observations Dataset, \\
    &&&&&& \url{https://doi.org/10.7283/T5J964RP}.\\\hline
    \multirow{2}{*}{P001} & \multirow{2}{*}{31.95} & \multirow{2}{*}{-112.80} & \multirow{2}{*}{GLM-07} & \multirow{2}{*}{GLM-22} & & UNAVCO Community, 2008, PBO GPS Network - P001-Organ\_PipeAZ2007 P.S., The GAGE Facility \\
    &&&&&& operated by UNAVCO, Inc., GPS/GNSS Observations Dataset, \url{https://doi.org/10.7283/T5DR2SGP}.\\\hline
